---
title: "ADS506 - Group Project - ERCOT Load Forecasting"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(astsa)
library(ggplot2)
library(readxl)
library(tidyverse)
library(ggdark)
library(reshape2) 
library(xts)
library(dplyr)
library(EnvStats)
```

Please see code blocks below which have been used to load, prep, transform and plot our data.

This will serve as our initial pre-processing, which we can further analyze for our modelling phase and our paper.

Further analysis will also be performed as we dive deeper into our paper and final code. 


```{r dataload}
load17.df <- read_excel('Native_Load_2017.xlsx')
load18.df <- read_excel('Native_Load_2018.xlsx')
load19.df <- read_excel('Native_Load_2019.xlsx')
load20.df <- read_excel('Native_Load_2020.xlsx')
load21.df <- read_excel('Native_Load_2021.xlsx')
```


```{r dataprep}
## The load dataset for 2017 had a space in the Timestamp column, preventing the binding of the four years
colnames(load17.df)[1] <- "HourEnding"

## Using bind_rows from tidyverse, the data from 2017-2020 was merged.
load_4_year <- bind_rows(load17.df,load18.df,load19.df,load20.df)
sum(is.na(load_4_year))

load_4_year <- load_4_year 



## Using strptime, the time variable data type was set. 
load_4_year$HourEnding <- as.POSIXct(strptime(load_4_year$HourEnding, format="%m/%d/%Y %H:%M"))

#Remove NA's which arise due to issues with daylight savings (1 record in each year)
load_4_year <- na.omit(load_4_year)

## Converting to a time series
ts_load <- xts(load_4_year, load_4_year$HourEnding)
ts_load <- ts_load[,-1]
head(ts_load)
```






```{r plotfunction, message=FALSE}
## A function has been created to plot data once appropriate transformations have been made. 

plot_data = function(df, title){
  
  plot_melt = melt(df, na.rm = FALSE, value.name = 'Load', id = 'HourEnding')
  names(plot_melt)[2] <- "Region"

  plt = ggplot(data = plot_melt, mapping = aes(x=HourEnding, y=Load, colour = Region)) +
  geom_line(size = 0.1, alpha=0.4) +
  ggdark::dark_mode() + 
  theme(legend.position = 'bottom', ) + 
  ggtitle(paste('ERCOT Actual Power Load by Region - ' , title)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Electrical Load (MW)") +
  xlab("Hour Ending") +
  theme(panel.grid.major = element_blank()) +
  guides(shape = guide_legend(override.aes = list(size = 1)),
               color = guide_legend(override.aes = list(size = 1))) +
  theme(legend.title = element_text(size = 8), 
              legend.text  = element_text(size = 8),
              legend.key.size = unit(1, "lines"))
    
  return(plt)}

plot_data(load_4_year, 'Raw Data')
```

```{r standardization, message=FALSE, warning=FALSE}

std = load_4_year %>%
  mutate_if(is.numeric, scale)

plot_data(std, 'Standardized')
```


```{r differencing, message=FALSE, warning=FALSE}
## Applying differencing and detrending to remove seasonality and make data stationary

dload <- data.frame(lapply(load_4_year[,-1], diff))
dload["HourEnding"] <- load_4_year$HourEnding[2:35060]

plot_data(dload, 'Differenced')
```


```{r Normalization, message=FALSE}
## Building out the min/max scaler to normalize the data as a function
normalize <- function(x) {
 (x-min(x))/(max(x)-min(x))
}

## Removing the POSIXlt column to normalize
loadnorm <- (load_4_year[,-1])

## Running the new DF through the normalization function
loadnorm <- normalize(loadnorm)

## Adding the POSIXlt column back on
loadnorm["HourEnding"] <- load_4_year$HourEnding

## Plotting the data
plot_data(loadnorm, "Normalized (Min/Max)")
```


It is clear that a min max normalization scalar is not a usable option as it does not account for the mean. The graph is still visually the same.


