---
title: "ADS506 - Group Project - ERCOT Load Forecasting"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(astsa)
library(ggplot2)
library(readxl)
library(tidyverse)
library(ggdark)
library(reshape2) 
library(xts)
library(dplyr)
library(EnvStats)
```

Please see code blocks below which have been used to load, prep, transform and plot our data.

This will serve as our initial pre-processing, which we can further analyze for our modelling phase and our paper.

Further analysis will also be performed as we dive deeper into our paper and final code. 


```{r dataload}
load17.df <- read_excel('Native_Load_2017.xlsx')
load18.df <- read_excel('Native_Load_2018.xlsx')
load19.df <- read_excel('Native_Load_2019.xlsx')
load20.df <- read_excel('Native_Load_2020.xlsx')
load21.df <- read_excel('Native_Load_2021.xlsx')
```


```{r dataprep}
## The load dataset for 2017 had a space in the Timestamp column, preventing the binding of the four years
colnames(load17.df)[1] <- "HourEnding"

## Using bind_rows from tidyverse, the data from 2017-2020 was merged.
load_4_year <- bind_rows(load17.df,load18.df,load19.df,load20.df)
sum(is.na(load_4_year))

load_4_year <- load_4_year[c("HourEnding","ERCOT")]



## Using strptime, the time variable data type was set. 
load_4_year$HourEnding <- as.POSIXct(strptime(load_4_year$HourEnding, format="%m/%d/%Y %H:%M"))

#Remove NA's which arise due to issues with daylight savings (1 record in each year)
load_4_year <- na.omit(load_4_year)

## Converting to a time series
ts_load <- xts(load_4_year, load_4_year$HourEnding)
ts_load <- ts_load[,-1]
head(ts_load)
```



It is clear that a min max normalization scalar is not a usable option as it does not account for the mean. The graph is still visually the same.


```{r bg_arima}

##### Bikram: notes here; im going to plot the P/ACF's, which should show if we need to difference, AR(x) and MA(x).

### From there, will fit the ARIMA model, forecast 2021 and then plot the dataset vs forecast



## Plot P/ACF of raw data



## Plot P/ACF of differenced data

```


